# (PART) Source file and output files {-}

# Rmd files {#rmd}

This chapter gathers general comments about `.Rmd` files.

## Options for all chuncks

It is convenient to set options for all the R chuncks of the document. This saves time when writing these chuncks.   
A natural place to set these options is in a first R chunck.


```{r, eval=FALSE}
knitr::opts_chunk$set(OPTION1 = TRUE/FALSE,
                      OPTION2 = TRUE/FALSE,
                      ...)
```


Importantly, these options are overriden by the particular chunck options.

````markdown
`r ''````{r, OPTION2=FALSE}
````




Options actually take R code. So, the following are examples that could be used to define the option.


````markdown
`r ''````{r, eval=4>3, echo=format(Sys.Date(), '%Y-%B-%d') > '2019-March-10'}
# eval is always TRUE
# echo = TRUE if current date is after March 10, 2019
```
````

The list of options can be found here [https://yihui.name/knitr/options/](https://yihui.name/knitr/options/). Below are some comments on some of these options (the least trivial for the author).

- `collapse` determines whether the source code and the ouput should be merged into a single block.
Here is the same chunck with different values of the option:

`collapse=TRUE`

```{r, collapse=TRUE}
2+ 2
3* 5
```

`collapse=FALSE`
    
```{r, collapse=FALSE}
2+ 2
3* 5
```

- `comment` gives the string to be printed before the output.

`comment='##'`

```{r, comment='##'}
2+ 2
```

`comment='R>'`
    
```{r, comment='R>'}
2+ 2
```

Worth noting: a `#` as a first character of the comment string (with `collpase=TRUE`) turns the output font into a comment-like text.

- `child` allows a document to call and use another file as input in the document.

````markdown
`r ''````{r, child='PATH/TO/OTHER/file.Rmd'}
````

The path can be either absolute or relative.  
For relative paths, the following applies:  

  - `~/` starts a path a the root,
  - `../` indicates the parent directory,
  - `../../` for parent of the parent directory,
  - to move forward, start with the name of the included folder in the current directory.



## Latex code

The overwhelming reason to introduce Latex code in a `.Rmd` file is for typesetting mathematical expressions.  
There are two main ways to type math in Latex:   

  - in the text, surrounded by special delimiters, `\( math \)` (alternatively, one can use the deprecated `$ math $`),
  - in an equation, surrounded by special delimiters, `\[ math equation \]`, or in a dedicated environment such as `\begin{equation} math equation \end{equation}` (also deprecated, `$$ math equation $$`).

[
Complete here with:  

  . examples of inline code and an equation,  
  . a reference for Latex,  
  . maybe the drawing interpreter.
]





