# (PART) Source and output files {-}

# Rmd files {#rmd}

This chapter gathers general comments about `.Rmd` files.

## Options for all chuncks

It is convenient to set options for all the R chuncks of the document. This saves time when writing these chuncks.   
A natural place to set these options is in a first R chunck.


```{r, eval=FALSE}
knitr::opts_chunk$set(OPTION1 = TRUE/FALSE,
                      OPTION2 = TRUE/FALSE,
                      ...)
```


Importantly, these options are overriden by the particular chunck options.

````markdown
`r ''````{r, OPTION2=FALSE}
````




Options actually take R code. So, the following are examples that could be used to define the option.


````markdown
`r ''````{r, eval=4>3, echo=format(Sys.Date(), '%Y-%B-%d') > '2019-March-10'}
# eval is always TRUE
# echo = TRUE if current date is after March 10, 2019
```
````

The list of options can be found here [https://yihui.name/knitr/options/](https://yihui.name/knitr/options/). Below are some comments on some of these options (the least trivial for the author).

- `collapse` determines whether the source code and the ouput should be merged into a single block.
Here is the same chunck with different values of the option:

`collapse=TRUE`

```{r, collapse=TRUE}
2+ 2
3* 5
```

`collapse=FALSE`
    
```{r, collapse=FALSE}
2+ 2
3* 5
```

- `comment` gives the string to be printed before the output.

`comment='##'`

```{r, comment='##'}
2+ 2
```

`comment='R>'`
    
```{r, comment='R>'}
2+ 2
```

Worth noting: a `#` as a first character of the comment string (with `collpase=TRUE`) turns the output font into a comment-like text.

- `child` allows a document to call and use another file as input in the document.

````markdown
`r ''````{r, child='PATH/TO/OTHER/file.Rmd'}
````

The path can be either absolute or relative.  
For relative paths, the following applies:  

  - `~/` starts a path a the root,
  - `../` indicates the parent directory,
  - `../../` for parent of the parent directory,
  - to move forward, start with the name of the included folder in the current directory.



## Latex code

The overwhelming reason to introduce Latex code in a `.Rmd` file is for typesetting mathematical expressions.  
There are two main ways to type math in Latex:   

  - in the text, surrounded by special delimiters, `\( math \)` (alternatively, one can use the deprecated `$ math $`),
  - in an equation, surrounded by special delimiters, `\[ math equation \]`, or in a dedicated environment such as `\begin{equation} math equation \end{equation}` (also deprecated, `$$ math equation $$`).

Math format is usually distinct from the text format. For instance, compare a^2^ + b^2^ = c^2^ (simple markdown) to \(a^2 + b^2 = c^2\) (Latex). Notice, however, that most mathematical symbols are not supported by markdown. This why basic knowledge of Latex is essential for producing documents with math formula.   
The first part of that knowledge consists in the list of Latex symbols. These are typed into a document starting with a slash `\`. For instance, the Latex symbol \(\alpha\) is produced by the code `\alpha` while \(\sum\) is given by `\sum`.   
Superscripts and subscripts are produced with the characters `^` and `_`, respectively, and the content of the super- and subscripts enclosed within `{ }`. For instance, \(x_{2}^{3}\) is produced by the code `x_{2}^{3}` (or `x^{3}_{2}`) and \(\sum_{i=1}^{4}\) is given by `\sum_{i=1}^{4}`.  
A comprehensive list of symbols (some of them requiring special Latex packages) can be found in @latexlist. Shorter lists can be found online, including on [wiki](https://oeis.org/wiki/List_of_LaTeX_mathematical_symbols).  
A web-based application helps beginners by offering suggestions to a hand-drawn symbol, [detexify](http://detexify.kirelabs.org/classify.html).    
A second bulk in the knowledge of Latex is the understanding of special environments. But then, if a document requires lots of special Latex environments, then `.Rmd` files are not the most appropriate type of files to use.  
Some of these environments, however, have a perfect integration in `.Rmd` files, such as the most common, the equation. Here is an example, given by the code

```markdown
\[
\sum_{i=0}^{\infty} ak^{i} = \frac{a}{1-k} \text{, for }  \lvert k \rvert < 1
\]
```
\[
\sum_{i=0}^{\infty} ak^{i} = \frac{a}{1-k} \text{, for }  \lvert k \rvert < 1
\]


## Include pictures

Some words about this topic.


## Include references

Some words about this topic.







