# Data stuctures {#datastructures}

This chapter lists the most common objects to store data.


## Atomic vectors
The basic data structure in R is the vector. Vectors have three characteristics:

  - a type, `typeof()`,
  - a length, `length()`,
  - some attributes, `attributes()`.

### Types of data 

There are four common types of atomic vectors:

  - logical,
  - integer,
  - double (often called numeric),
  - character.

Note that if vector elements not all of the same type, R makes coercion.  In that case, the order becomes:
"logical < numeric < character".

Here are a few illustrations.

```{r}
trop1 <- c(24, 23, 23, 24, 24)
trop1

trop2 <- c("mi", "ga", "tr", "ba", "ha")
trop2

trop3 <- c(24, 24, "trierweiler")
trop3
typeof(trop3)

var <- c(2)
var

tar <- c("2")
tar

var + 30
#tar + 30
```



```{r}
log_vector <- c(TRUE, FALSE, FALSE)

int_vector <- c(12, 10, 3, "tre")
typeof(int_vector)
int_vector <- c(12L, 10L, 3L)
typeof(int_vector)

num_vector <- c(12, 10, 3)
typeof(num_vector)
chr_vector <- c("a", "b", "c")
typeof(chr_vector)
is.numeric(num_vector)

no_vector <- c(2, "a")
is.numeric(no_vector)
typeof(no_vector)
```



```{r vectors}
vector_A <- c(1:5)
vector_A
length(vector_A)
```

### Factor vector

This is a special type of vector. It has a limited number of values, called levels. These levels can be unordered (e.g., gender is either "Female" or "Male") or ordered (e.g. school level is "Primary", "Secondary", "Tertiary")
```{r factor}
gender <-factor(c("Male", "Male", "Female", "Male", "Female", "Female", "Male"))
gender
levels(gender)
summary(gender)
school<-factor(c("Primary", "Secondary", "Tertiary"), ordered=TRUE)
school

school2<-factor(c("Primary", "Secondary","Secondary", "Tertiary"), labels=c( "Secondary", "Tertiary","Primary"), ordered=TRUE)
school2

```

## Matrices and arrays
My experience is that R is not often used for matrices calculations: it is too slow for that, that are better programs for that out there (e.g. Matlab).

```{r}
M <- matrix(c(4, 1, 0, 3, 6, 8), nrow=3, ncol=2) 
M
```

If we think of a matrix as a 2 dimensions vector, then arrays are $n$ dimensions vectors. Is it important? Probably in some cases, not so much for us.

```{r array example}
mya<-array(data=1:18, dim=c(2,3,3))
mya
```

## Lists
These are the one-size fit all structure... A list is an object composed of any other object, even... another list! Very useful data structure!

```{r list example}
school<-factor(c("Primary", "Secondary", "Tertiary"), ordered=TRUE)
mylist <- list(numbers=c(1:60), somenames=c("Jim","Jules"), results= c(T,F,F,T), school=school)
mylist
names(mylist) <- c("N", "O","R","S")
```

## Data frames
The second most important data structure in R. You can think of it as a better version of a data set in Excel. It stacks together observations over many variables, each of these variables being a vector.

```{r dataframe}
data(mtcars)
class(mtcars)
mtcars
head(mtcars)
str(mtcars)
names(mtcars) # names of the variables in the data frame

length(mtcars)

str(mtcars)
```