# Functions {#functions}

Everything that happens in R is a function call. If we want R to do something, we must use a function, may it be as simple as a parentheses `(`. 
If the function is not built-in or coming in a package, one must write it.
A function is itself an object. When applied to another object, the function "makes something" to that object and returns an object.

## Structure of a simple function

Here is a simple example of a function that returns the square of the object provided.

```{r}
pow_two <- function(a){
  a^2
}
pow_two(12)

```



### Name of the function
In our example above, the name was `pow_two`. To call, i.e., to use the function later in the code, type `pow_two()`.

### Arguments

The arguments are given in the parentheses right after `function`.  
Here, we use an argument `a`, which can be any object, e.g., a vector
```{r}
pow_two(1:10)
```
When using this function, we can see that one argument only must be passed.  
A value for the argument is then given in the parentheses when calling the function, e.g., `pow_two(12)`.  
A function can have several arguments

### Commands
The commands will be executed, if possible.  
One command source of error is when the class of the argument provided is NOT compatible with the commands in the function

```{r , error=TRUE}
pow_two("hello") # this can't work
```


### Return of a function

The return of a function is given by the last expression executed inside the function.

```{r }
my_f <- function(a){
  a^2
  a^3 # THIS is the last expression executed
}
my_f(5)
```

We can use `return()` in the function in order to chose what it  returns.

```{r }
my_f2 <- function(a){
  return(a^2)
  a^3
}
my_f2(5)
```

A function can return only one object; this is not really a  problem as we  can always stack everything in a `list()` and have the function return the list.

```{r }
my_f3 <- function(a){
  list(ptwo=a^2, pthree=a^3, pfour=a^4) 
}
my_f3(5:20)
```

If we want to use the return of our function, we must assign it to an object.

```{r }
x <- my_f3(1:5) # evaluates the function above on the vector 1:5
x$pthree # shows the element pthree of the saved list x
```

An example where we better use `return` is the following

```{r}
pow_three <- function(a){
  if (!is.numeric(a)) {
    return(print("You must give a numeric vector!"))
  } else {
    return(a^3)
  }
  
}
pow_three(1:4)
```



## Multiple arguments and their identification

As mentioned above, a function can have several arguments.

```{r , error= TRUE}
my_f4 <- function(abc, xyz){
  c(abc^2, xyz^3)
}
my_f4(5) # this won't work because xyz is missing
my_f4(5, 10)
```

R has at least two ways of identifying arguments, by position and by name.  
If no name is specified, R uses the arguments according to their positions in the definition of the function.  
In the example above, R understands that `abc=5` and `xyz=10` because of the way they are given in the call `my_f4(5, 10)`.  
Identification by names would be
```{r , error= TRUE}
my_f4(abc=5, xyz=10) # same as before
my_f4(xyz=5, abc=10) 
# different from before because the names are more important than the position
my_f4(xyz=5, 10) # a mix between name and position 
```

### Default values

A function can be defined with default values for their arguments.  
R will use the values provided in the call; if one is missing, R will use the default value.

```{r , error= TRUE}

my_f4 <- function(abc, xyz=2){
  c(abc^2, xyz^3)
}
my_f4(5) # no second argument? doesn't matter, the function has a 
# default value for it, i.e., 2

my_f5 <- function(a, tada){
  if (tada==TRUE){
    plot(a)
  } else {
    sum(a)
  }
}


my_f5(1:20, TRUE)

my_f6 <- function(a, tada=TRUE){
  if (tada==TRUE){
    plot(a)
  } else {
    sum(a)
  }
}
my_f6(1:30)
```


## Scoping
  
A function creates its own environment. In turn, this environment of the function is a child of the environment where the function was created.  
Scoping refers to the set of rules on what environment R looks for the value of a name. For instance, if the code contains `myv <- 12`, then later typing `myv` forces R to retrieve 12 from an environment.  
R uses lexical scoping and, under some circumstances, dynamic scoping. We're interested in lexical scoping.


```{r}
mean <- function(x) {
  x+1
}
mean(1:10)

rm(mean)

mean(1:10)
```


### Rule 1: name masking  

If a name is not defined in a function, R looks one level up, i.e., in the parent:

```{r}
a <- 10

f <- function(){
  b <- 2
  a + b # same as return(a + b)
}

f()

c <- 11
g <- function() {
  d <- 4
  h <- function() {
    e <- 5
    c + 2*d + e
  }
  h()
}

g()



c <- 11
d <- 100
g <- function() {
  d <- 4
  h <- function() {
    e <- 5
    c + 2*d + e
  }
  h()
}
 g()
  
 
```



### Rule 2: fresh start

Every time a function is called, a new environment is created. In other words, the function does not know that it has been called before.

```{r}
a 
t <- function() {
  if (exists("a")) { 
    a <- a + 2
    a
  } else {
    a <- 20 
    a
  } 
}
t()

t()
t()

a <- 4
# gives 6 because when the function t looks for 'a', NOW, it finds it in the parent 
t() 


l <- function() {
  a <- 5 
  # 'a' is defined in the environment, so there is no need to look into the parent 
  if (exists("a")) { 
    a <- a + 2
    a
  } else {
    a <- 10 
    a
  } 
}
l()

```

### Rule 3: dynamic lookup

R looks for values when the function is called. If the environment where R will look for values has changed in between calls, the result of the function may be different. This is a source of error that should avoided.

```{r}
u <- function() a^2
a <- 2
u() 
# needs a value for 'a' because it does not define one within its environment; 
# therefore, u() looks up to, in this case, the global environment
a <- 3
u()
```

### Functions inside functions

Consider the following example of a function, it returns a function!


```{r}
m.power <- function(pow){
  m.exp <- function(b){ b^pow }
  m.exp # return the m.exp function
}
```

We can create many functions with that function. Importantly, notice that functions preserve the environment in which they were created.

```{r}

mycube <- m.power(3)
mycube(4)

mysquare <- m.power(2)
mysquare(5)

# ls(environment(m.power)) gives the global environment
ls(environment(mycube)) 

w <- function(k){
  q <- 10
  k*q
}

z <- function() w(2) # z is itself a function that inherits the environment of w
# but z, here, is forced to use 2 as an argument 

z()
```


### Examples


The following example illustrates lexical scoping.
```{r}
yy <- 10

ff <- function(xx){
  yy <- 2
  yy^2 + gg(xx)
}

gg <- function(xx){
  xx*yy
}

ff(3) 
```


As another illustration, we could try to guess what  `g(2)` produces after the following commands. Hint: remember the in which environment the functions were created because it tells you what are the correct bindings (parent).


```{r}
a <- 1
b <- 2
f <- function(x){
  a*x + b
}
g <- function(x){
  a <- 2
  b <- 1
  f(x)
}

g(2)
```

Compare to `g(2)` here.

```{r}
a <- 1
b <- 2
f <- function(a,b){
  return( function(x) {a*x + b})
}
g <- f(2,1)
g(2)
```















