---
title: "DS4B: Exercises"
author: "A. Fidalgo"
date: '**Last update:** `r format(Sys.time(), "%B %d, %H:%M")`'
output:
  rmdformats::readthedown:
    css: custom_readthedown.css
    number_sections: no
    highlight: tango
  pdf_document:
    includes:
      in_header: 'preamble.tex'
    highlight: "tango"
    number_sections: yes
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following are a question sets to help master the concepts that we discussed.

# Lecture on March 22

## Global chunk options

We can use options, e.g., `echo`, to control the output and the format of the R chunks.  
Suppose you want to use the same options for all the R chuncks. It would be tedious to write it _n_ times!    
How would you set that option once for the whole document? (Hint: not a YAML issue.)    
How would you make an exception for one specific R chunk?


## Yet another output format for the same `Rmd` file

The YAML of this `Rmd` file contains the following code.

```{r, eval=FALSE}
output:
  prettydoc::html_pretty:  
    theme: architect  
    highlight: github
  pdf_document:
    highlight: tango
    number_sections: no
    toc: no  
```

If you `Knit` the document as such, it will almost certainly produce an error. Why?   
Make sure you can `Knit` this `Rmd` file with **all** the output options.

## Dynamic document 1.0

There is a dynamic element in the title part of this document.   
Guess how it works and modify it at your will.

## Line break in `Rmd` file

The text in the `Rmd` file looks exactly like copied below as code.
Why doesn't the output file (whatever its type) reflect the line break?

```{r, eval=FALSE, echo= TRUE}
The text in the `Rmd` file looks exactly like copied below as code.
Why doesn't the output file (whatever its type) reflect the line break?
```



















  
    
      
        
        



