---
title: "Ongoing Tasks"
author: "A. Fidalgo"
date: '**Last update:** `r format(Sys.time(), "%B %d, %H:%m")`'
output:
  pdf_document:
    highlight: "tango"
    number_sections: yes
    toc: yes
  html_document:
    toc: no
  html_notebook:
    code_folding: hide
    highlight: tango
    number_sections: no
    theme: cerulean
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  word_document:
    toc: no
    toc_depth: '2'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

--- 
This document is thread of the tasks that structure our class. Of course, not all of them will be documented here. It would be impossible to keep such a level of detail. But those below are certainly compulsory benchmarks.  
Notice that the file will be updated in reverse chronological order (last section entries appear first; normal order _within_ the section). That is meant to help follow the thread. The table of contents, specially on HTML-like formats, will be very useful to navigate the tasks.  


# Homework for March 22

## Train the workflow

Play around with the example that we did in class to illustrate **reproducible research** (rr) and **dynamic documents** (dd).   
The file is on GitHub, but I slightly modified the YAML in order to allow the illustration of cross-references of sections. This new format (`bookdown::pdf_document2`) requires a R package. Hence, you need to install it with `install.packages('bookdown')` or by using the menu _Tools> Install Packages..._.



# Lecture on March 15 

## Check your R/Rstudio/Latex installation

Create a new Rmd file _File> New File> R Markdown_. And then click on Knitr, possibly by chosing the output format on the scroll down menu (i.e., Knit to...).  


If needed, install required packages (e.g., `knitr`).

```{r, eval=FALSE}
install.packages("knitr")
```

Notice that, in order to create a pdf document, you must have a Latex distribution installed; similarly for Word output, Microsoft Word is necessary.

## Create own Rmd file

Notice that everytime such new file is created with the menu, it comes with  pre-populated content. It can help, but it can also be annoying because you need first to erase it and start writing your stuff.  

Your first Rmd can be a change in this default content or a copy from the class GitHub (or your very own).


## Create class folders

In your machine, create a folder for this class. Within that folder, create a folder for your book. Members of the same group must have the **same name** for that book folder.   
Put your first Rmd file into your book folder.



## Create your project

_File> New project> Existing Directory_ and chose that book folder.  
Now, everytime you create content for your book, you must start a Rstudio session _File> Open Project..._  
All the files of the project are the files of the folder, and vice versa.




# Prior to the first Lecture 

## Install applications

Please download the following FREE applications, available on all platforms:

  1. [R (https://cran.uni-muenster.de/)](https://cran.uni-muenster.de/) 
  2. [RStudio, free Desktop version (https://www.rstudio.com/products/rstudio/download/#download)](https://www.rstudio.com/products/rstudio/download/#download)
  3. [A Latex distribution](https://www.latex-project.org/get/) (e.g., MacTex for Mac or MiKTeX for Windows machines) 

The first two are easily and quickly installed. The last is a very large program (a few Gb) and needs time to install.




# Planned tasks {-}

This is a very incomplete list of tasks that will enter our workflow at some point. They are mentioned here for a reference.

## Install a Git application {-}

Another application needed in the class is a [Git](https://git-scm.com/downloads) distribution. This is also a free software.

## Set up GitHub {-}

One member of the group must create a free [GitHub](https://github.com/join) account. In that account, you will create a new repository (pronounced 'repo') whose name is exactly the same as your R project / book folder (e.g., 'myRbook').  
The linking between GitHub and Rstudio will be explained later.

## Include images {-}

## Referencing of sections/figures/equations... {-}

## Introduction to Latex {-}

This is particularly important for writing math expressions.






