---
title: "DS4B: Ongoing Tasks"
author: "A. Fidalgo"
date: '**Last update:** `r format(Sys.time(), "%B %d, %H:%M")`'
output:
  pdf_document:
    includes:
      in_header: 'preamble.tex'
    highlight: "tango"
    number_sections: yes
    toc: no
  rmdformats::readthedown:
    css: custom_readthedown.css
    number_sections: no
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



This document is a thread of the tasks that structure our class. Of course, not all of them will be documented here. It would be impossible to keep such a level of detail. But those below are certainly compulsory benchmarks.  
Notice that the file will be updated in reverse chronological order (last section entries appear first; normal order _within_ the section). That is meant to help follow the thread. The table of contents, specially on HTML-like formats, will be very useful to navigate the tasks.  
The pdf file (and only it) is instructed to count the sections in reverse order.  
A preview of the HTML file, which is easier to navigate, is available: 
 [preview html here.](https://htmlpreview.github.io/?https://github.com/AntonioFidalgo/MIBM/blob/master/tasks.html)



# Lecture on March 29

## Integrate the material

On the topics (see notes):

  - data structures,
  - simple functions,
  - sub-setting,
  - conditions,
  - functions (time allowing).

## Exercises

Check out the `Exercises.pdf` (or other format) file and answer the current set of exercises on:

  - simple vectors creation,
  - sub-setting,
  - functions (time allowing). 

# Homework for March 29


## Exercises 

Check out the `Exercises.pdf` (or other format) file and answer the current set of exercises on 



## Set up GitHub for team collaboration

Access the chapter in my book and follow the instructions for setting up GitHub for team collaboration.  
Importantly, this must be done **only once** you have created your own experimental repo on GitHub and have managed to push/pull code to/from it to your machine.



# Lecture on March 22

## Build minimal book on your machine

Install the required package, if you haven't done it: `install.packages("bookdown")`.  
Download the minimal bookdown example at [https://github.com/rstudio/bookdown-demo](https://github.com/rstudio/bookdown-demo).  
Copy all the files of that folder into your book folder.  
Change at least the three following files with self-describing fields (e.g., author):  

  . `_output.yml`  
  . `_bookdown.yml`  
  . `index.Rmd`

Build the book in different formats.  
That's all: you are ready to write your book.

## Understand the role of the files in the folder

Analyze / change / add / remove files from your folder in order to understand how to structure a book with Rmd files. Check the following points (non-exhaustive list):  

  . what files are included in the output,   
  . how these files are structured and how they enter the book,   
  . what is the role of the other files in the folder.

Get help at the book about bookdown, which is a book written with bookdown at [https://bookdown.org/yihui/bookdown/](https://bookdown.org/yihui/bookdown/).



## Install a Git application

Another application needed in the class is a [Git (https://git-scm.com/downloads)](https://git-scm.com/downloads) distribution. This is also a free software.  
Once you have installed Git for your version control, activate it in RStudio: _Tools> Global Options> Git/SVN_ and click on _Enable version control interface for RStudio projects_.  
Also generate a SHH RSA key. We will use it to identify at the GitHub repo.

## Sign for GitHub

Create an account at GitHub at [https://github.com](https://github.com/join).


## Create GitHub repo and link your machine to it

One member of each group must create a new repository (pronounced 'repo') whose name is **exactly** the same as your R project / book folder (e.g., 'myRbook').   
On the top left menu in GitHub, go to _Settings> SHH and GPG keys_ and click on 'New SHH key'. Paste the SHH key generated by RStudio.

In RStudio go to _Tools> Project Options...> Git/SVN_. Under _Version control system_, select 'Git'.    
Still in RStudio, _Tools> Terminal> New Terminal_. This open a Terminal where you can paste the message shown at the creation of the repo (changing the names, of course):

```{r, eval=FALSE}
git remote add origin https://github.com/YOURNAME/YOURREPO.git
git push -u origin master
```

Your local `master` should now be connected to the `master` on GitHub.  
If necessary, restart RStudio. At the restart, a Git thumbnail should appear in a pane. You are ready to commit and push your files.


## Structure of the book

Make a first attempt at designing a structure for your book. Create empty chapters as placeholders.


## First steps in R

Integrate the chapter on R as a calculator and the chapter on data structures in R.




# Homework for March 22

## Train the workflow

Play around with the example that we did in class to illustrate **reproducible research** (rr) and **dynamic documents** (dd).   
The file is on GitHub, but I slightly modified the YAML in order to allow the illustration of cross-references of sections. This new format (`bookdown::pdf_document2`) requires a R package. Hence, you need to install it with `install.packages('bookdown')` or by using the menu _Tools> Install Packages..._



# Lecture on March 15 

## Check your R/Rstudio/Latex installation

Create a new Rmd file _File> New File> R Markdown_. And then click on Knitr, possibly by choosing the output format on the scroll down menu (i.e., Knit to...).  


If needed, install required packages (e.g., `knitr`).

```{r, eval=FALSE}
install.packages("knitr")
```

Notice that, in order to create a pdf document, you must have a Latex distribution installed; similarly for Word output, Microsoft Word is necessary.

## Create own Rmd file

Notice that every time such new file is created with the menu, it comes with  pre-populated content. It can help, but it can also be annoying because you need first to erase it and start writing your stuff.  

Your first Rmd can be a change in this default content or a copy from the class GitHub (or your very own).


## Create class folders

In your machine, create a folder for this class. Within that folder, create a folder for your book. Members of the same group must have the **same name** for that book folder.   
Put your first Rmd file into your book folder.



## Create your project

_File> New project> Existing Directory_ and chose that book folder.  
Now, every time you create content for your book, you must start a Rstudio session _File> Open Project..._  
All the files of the project are the files of the folder, and vice versa.




# Prior to the first Lecture 

## Install applications

Please download the following FREE applications, available on all platforms:

  1. [R (https://cran.uni-muenster.de/)](https://cran.uni-muenster.de/) 
  2. [RStudio, free Desktop version (https://www.rstudio.com/products/rstudio/download/#download)](https://www.rstudio.com/products/rstudio/download/#download)
  3. [A Latex distribution](https://www.latex-project.org/get/) (e.g., MacTex for Mac or MiKTeX for Windows machines) 

The first two are easily and quickly installed. The last is a very large program (a few Gb) and needs time to install.




# Planned tasks {-}

This is a very incomplete list of tasks that will enter our workflow at some point. They are mentioned here for a reference.



## Include images {-}

## Referencing of sections/figures/equations... {-}

## Introduction to Latex {-}

This is particularly important for writing math expressions and for output customization.

## Introduction to html customization (css) {-}

As requested by Mr. Miebach.

## Easy slides in RMarkdow {-}

As requested by Mr. Gaulke.


























  
    
      
        
        



